---
title: "Data Analysis Project"
author: "Mauriah Allen"
date: "4-23-24"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are checking out the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data
wbb_2022 <- read_csv("data/wbb_rosters_2022_23.csv") |> clean_names() 

wbb_2023 <-read_csv("data/wbb_rosters_2023_24.csv")|> clean_names() 

wbb_2021 <-read_csv("data/wbb_rosters_2021_22.csv")|> clean_names()

wbb_2020 <- read_csv("data/wbb_rosters_2020_21.csv")|> clean_names()

wbb_crosswalk <- read_csv("data/wbb_rosters23_crosswalk.csv") |> clean_names()

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: What is the average height of the athletes in each year from the rosters 2020-2023, which year had the tallest average
-   **Analysis summary**: #so far 2020 has the highest average height so far at the years with the height of 6.8

```{r}
# Put code to reverse engineer sentence here
wbb_2020_height_per_town <- wbb_2020 |> 
group_by(hometown_clean) |> summarise(Avg_height_per_town = mean(total_inches)) |> arrange(desc(Avg_height_per_town)) |> mutate(feet = Avg_height_per_town/12)

#still needs 2021, because we need to mutate a coloumn for the total inches. 


wbb_2022_height_per_town <- wbb_2022 |> 
group_by(hometown) |> summarise(Avg_height_per_town = mean(total_inches)) |> arrange(desc(Avg_height_per_town)) |> mutate(feet = Avg_height_per_town/12)

wbb_2023_height_per_town <- wbb_2023 |> 
group_by(hometown) |> summarise(Avg_height_per_town = mean(total_inches)) |> arrange(desc(Avg_height_per_town)) |> mutate(feet = Avg_height_per_town/12)


# Display results of code below this codeblock

```

### Sentence 2

-   **Sentence text**: How does the quanity of the position of a Forward change over the years?
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock
wbb_positions_2020 <- wbb_2020 |> 
    group_by(position_clean) |> 
    summarise(count = n())

wbb_positions_2021 <- wbb_2021 |> 
    group_by(position_clean) |> 
    summarise(count = n())

wbb_positions_2022 <- wbb_2022 |> 
    group_by(position_clean) |> 
    summarise(count = n())

wbb_positions_2023 <- wbb_2023 |> 
    group_by(position_clean) |> 
    summarise(count = n())

```

### Sentence 3

-   **Sentence text**: In the years 2022 and 2023, how many teams had more than 1 redshirt on their team? What team had the most, and where are they from?
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
redshirt_22 <- wbb_2022 |>
  group_by(team, team_state) |> 
  summarise(redshirt_numbers = sum(redshirt)) |> arrange(desc(redshirt_numbers)) |> 
  filter(redshirt_numbers > 1) 

redshirt_23 <- wbb_2023 |>
  group_by(team, team_state) |> 
  summarise(redshirt_count = sum(redshirt)) |> arrange(desc(redshirt_count)) |> 
  filter(redshirt_count > 1)

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: How many players are from DMV each year, does the number increase or decrease or fluncuates?
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
